\documentclass{../weeklyslides}

\addbibresource{../../References.bib}

\title[Weekly Report]{ECE 497: Special Project \\ Weekly Report}
\subtitle{Week 09}
\author{Alexander Lukens \and Karl Hallsby}
\institute{Illinois Institute of Technology}
\date{\DTMdisplaydate{2021}{3}{25}{-1}}

\begin{document}

\nocite{chipyard}

\begin{frame}
  \titlepage{}
\end{frame}

\section{What We Did}
\begin{frame}
  \frametitle{What We Did}
  \begin{itemize}
	\item Successfully implemented Chipyard design with ability to output over UART
	\item Was solved by combination of Chipyard dev doing work on a branch \& us solving an issue with the general FPGA implementation
	\item Error in the previous Uart "HarnessBinder" that attached the UART interface to the FPGA. Found by inspection
  \end{itemize}
\end{frame}

\begin{frame}
	\frametitle{What We Did}
  \begin{itemize}
	\item Solved erroneous UART characters \& glitches. Use command from Freedom-e-sdk: \mintinline{bash}{metal_tty_putc()} to output char[] to serial interface
	
	\item Made simple monitor program to receive user input. Working on parsing commands in C
	
	\item Reformatted documentation tex file to improve reading \& writing abilities. Writing source files now simpler, reading output pdf simpler. Useful for final report, documentation, and ECE day presentation
	
	\item Elaborated on the setup process in our documentation with specific step-by-step instructions.
  \end{itemize}
\end{frame}
\section{Demo}\label{sec:Demo}
\begin{frame}
	\frametitle{Demo}
	\begin{itemize}
		\item Bitstream upload to FPGA
		\item Flash C Program to FPGA memory
		\item Monitor program in serial terminal
	\end{itemize}
	
\end{frame}

\section{What We Learned}\label{sec:What_We_Learned}
\begin{frame}
  \frametitle{What We Learned}
  \begin{itemize}
  	\item There were several issues preventing UART communication on the FPGA. Improper implementation
  	\item Freedom-e-sdk does not instantiate GPIO correctly $\rightarrow$ need to manually edit GPIO to make buttons/switches/leds work in our C code.
  	\item Writing comprehensive documentation is time consuming
  \end{itemize}
\end{frame}

\section{Next Steps}\label{sec:Next_Steps}
\begin{frame}
  \frametitle{Next Steps}
  \begin{itemize}
  \item Prepare full documentation of what we have learned so far.
    \begin{itemize}
    \item Debugger $\rightarrow$ Setup, Upload, Tethered debug
    \item Chipyard and its build process $\rightarrow$ setup, examples, supplementatal projects
    \item Software used
    \item Monitor Program for FPGA. Functionality to include memory examination, execute program at specific address, IO tests
    \end{itemize}
  \item Make recorded presentation for ECE Research Day
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{References}
  
  \printbibliography[heading=bibintoc]{}
\end{frame}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
