\documentclass{../weeklyslides}

\addbibresource{../../References.bib}

\title[Weekly Report]{ECE 497: Special Project \\ Weekly Report}
\subtitle{Week 07}
\author{Alexander Lukens \and Karl Hallsby}
\institute{Illinois Institute of Technology}
\date{\DTMdisplaydate{2021}{3}{11}{-1}}

\begin{document}

\nocite{chipyard}

\begin{frame}
  \titlepage{}
\end{frame}

\section{What We Did}\label{sec:What_We_Did}
\begin{frame}
  \frametitle{\nameref*{sec:What_We_Did}}
  \begin{itemize}
  \item Attempted to implement UART loopback interface in Verilog and flash to FPGA.\@
    \begin{itemize}
    \item Worked correctly immediately, and on first try.
    \item Means that
    \end{itemize}
  \item Karl received a response from ChipYard maintainer, who redirected us to a Google Group.
  \item Both of us were preparing for exams, so we did not have as much progress.
    \begin{itemize}
    \item Investigating using older Vivado versions for writing bitstream
    \item Still working on the submodule thing.
    \end{itemize}
  \end{itemize}
\end{frame}

\section{What We Learned}\label{sec:What_We_Learned}
\begin{frame}
  \frametitle{\nameref*{sec:What_We_Learned}}
  \begin{itemize}
  \item ChipYard FPGA image doesn't correctly do something with UART communication.
  \item ChipYard has a Google Group for getting help/support with ChipYard.
    \begin{itemize}
    \item Need to interact with them to narrow down our possible problems.
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Next Steps}\label{sec:Next_Steps}
\begin{frame}
  \frametitle{\nameref*{sec:Next_Steps}}
  \begin{itemize}
  \item Prepare full documentation of what we have learned so far.
  \item Identify reasonable goals in preparation for ECE Research Day.
  \item Read other Verilog simulators to have a better grasp of ChipYard's.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{References}
  \printbibliography[heading=bibintoc]{}
\end{frame}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
